InnoDB's Full Text Search overview (with Shiny/R)

### 2016-04-20 article.rmd
Final changes

```{r init, message = FALSE, echo=FALSE}
library(png)
#options(width = 9999)
#knitr::opts_chunk$set(fig.width=9999)
# opts_chunk$set(fig.width=9999) #, fig.height=6))

library(knitr)
library(shiny)
require(RMySQL)
library(ggplot2)

# Comment if you want to avoid reload all the data
# If you want to start from scratch, uncomment the following line.
# Otherwise, you can restore the dumps located in the dumps folder
#source('~/fts_article/load.R')

@@

# You do not want clear passwords, but is a sandbox.
con <- dbConnect(MySQL(),
                 user="msandbox", password="msandbox", port=13454
                 ,dbname="test", socket="/tmp/mysql_sandbox13454.sock",
                 host="127.0.0.1")

#con <- dbConnect(MySQL(),
#                 user="ftsinnodb", password="elfilete", port=3306
#                 ,dbname="ftsinnodb", #socket="/tmp/mysql_sandbox14901.sock",
#                 host="db4free.net")

@@

Repository available at [Github](https://github.com/3manuek/fts_article).

@@

@@

The point of this "article" is to show dynamically how the search algorithms work,
using non-fictional data (data sources were downloaded from www.gutenberg.org
project) within an easy interface.

@@

The current article has been developed using Shiny/R in order to allow you to see
the effects of the algorithms. It connects to a open database at `db4free.net`,
which unfortunately is not 100% available.

@@

@@

```{r, echo=FALSE}

sql <- "SELECT bp.bookid, round(avg(length(bp.content))) as avgLength, round(STD(length(bp.content))) as stdLength, count(*) as numEntries
            FROM bookContent bp  
            GROUP BY bookid;"
rs <- dbGetQuery(con, sql)
return(rs)

bookContentStats <- rs

sql <- "SELECT bl.bookid, round(avg(length(bl.content))) as avgLength, round(STD(length(bl.content))) as stdLength, count(*) as numEntries
            FROM  bookContentByLine bl
            GROUP BY bookid;"
rs <- dbGetQuery(con, sql)
return(rs)

bookContentByLineStats <- rs

@@

### Bug

Submitted https://bugs.mysql.com/bug.php?id=80347

```
emanuel@3laptop ~/sandboxes/rsandbox_5_7_9 $ ./m dumpTest < full.dump 
ERROR 182 (HY000) at line 73: Invalid InnoDB FTS Doc ID
```

It works if loaded ddl and data separately.

```
emanuel@3laptop ~/sandboxes/rsandbox_5_7_9 $ ./m dumpTest < ddl.dump 
emanuel@3laptop ~/sandboxes/rsandbox_5_7_9 $ ./m dumpTest < onlyData.dump 
emanuel@3laptop ~/sandboxes/rsandbox_5_7_9 $ ./m dumpTest < full.dump 
ERROR 182 (HY000) at line 73: Invalid InnoDB FTS Doc ID
```

Changed innodb_autoinc_lock_mode to 0 (old behavior) and 2, with the same results.
Looks like FTS in InnoDB does not like metadata locks (as we are using a single transaction
to load the data).

>>>>

```{r init, message = FALSE, echo=FALSE,include=FALSE}
library(png)
#options(width = 9999)
#knitr::opts_chunk$set(fig.width=9999)
# opts_chunk$set(fig.width=9999) #, fig.height=6))

library(knitr)
library(shiny)
suppressWarnings(library(RMySQL))
library(ggplot2)

# Comment if you want to avoid reload all the data
# If you want to start from scratch, uncomment the following line.
# Otherwise, you can restore the dumps located in the dumps folder

#source('~/fts_article/load.R')

mypass <- scan(file = '.key', what = (pass=""))
RDShost <- scan(file = '.RDShost', what = (RDShost=""))

@@

# You do not want clear passwords, but is a sandbox.
con <- dbConnect(MySQL(),
                 user="mysql57", password=mypass, port=3306
                 ,dbname="ftslab", #socket="/tmp/mysql_sandbox13454.sock",
                 host=RDShost)

@@

@@

> Repository available at [Github](https://github.com/3manuek/fts_article).

@@

The point of this _article_ is to show dynamically how the search algorithms work,
using non-fictional data (data sources were downloaded from www.gutenberg.org
project) within an easy interface.

@@

The current article has been developed using Shiny/R in order to allow you to see
the effects of the algorithms.

@@

## Why I consider FTS sometimes misunderstood?

FTS is a technology that can be use for any purpose, not only simple searches.
There is a myth that the FTS only should be placed on clusters for that purpose,
which I agree. However, certain bussiness rules require comple searches, and 
having such feature can be a win.

RDBMS aren't a good place for massive amount of FTS queries, without using any
of the join capabilities that they offer, or the ACID complaints.

As I said above, FTS is totally acceptable in RDBMS, if you are using at least
one RDBMS critical feature, required by your bussines model.

@@

```{r, ignore=FALSE, echo=FALSE, message=FALSE}

sql <- "SELECT bp.bookid, round(avg(length(bp.content))) as avgLength, round(STD(length(bp.content))) as stdLength, count(*) as numEntries
            FROM bookContent bp  
            GROUP BY bookid;"
rs1 <- dbGetQuery(con, sql)

bookContentStats <- rs1

sql <- "SELECT bl.bookid, round(avg(length(bl.content))) as avgLength, round(STD(length(bl.content))) as stdLength, count(*) as numEntries
            FROM  bookContentByLine bl
            GROUP BY bookid;"
rs2 <- dbGetQuery(con, sql)

bookContentByLineStats <- rs2

```{r, echo=FALSE}
return(rs1)
return(rs2)

@@

### Bug 80347 - Invalid InnoDB FTS Doc ID

> Submitted https://bugs.mysql.com/bug.php?id=80347

```
emanuel@3laptop ~/sandboxes/rsandbox_5_7_9 $ ./m dumpTest < full.dump 
ERROR 182 (HY000) at line 73: Invalid InnoDB FTS Doc ID

emanuel@3laptop ~/sandboxes/rsandbox_5_7_9 $ ./m dumpTest < ddl.dump 
emanuel@3laptop ~/sandboxes/rsandbox_5_7_9 $ ./m dumpTest < onlyData.dump 
emanuel@3laptop ~/sandboxes/rsandbox_5_7_9 $ ./m dumpTest < full.dump 
ERROR 182 (HY000) at line 73: Invalid InnoDB FTS Doc ID
```

mysqldump is not very clever if you use `FTS_DOC_ID`:

```
2016-02-13T22:11:53.125300Z 19 [ERROR] InnoDB: Doc ID 10002 is too big. Its difference with largest used Doc ID 1 cannot exceed or equal to 10000
```

It takes dumps without considering the restriction coded in 
`innobase/row/row0mysql.cc`:

```
Difference between Doc IDs are restricted within
4 bytes integer. See fts_get_encoded_len() 
```

The fix to this is backuping the table by chunks of 10000 documents.
